---
title: "p8105_hw3_qs2261"
author: "Qinting Shen"
date: "2023-10-14"
output: github_document
---


```{r}
library(tidyverse)
```

### Problem 1

load the dataset `instacart`
```{r}
library(p8105.datasets)
data("instacart")
```

`instacart` has `r nrow(instacart)` observations and `r ncol(instacart)` variables. This dataset collects some information of orders from instacart, including `order_id`, `user_id`, `order_hour_of_day` `days_since_prior_order`, `product_name`, `aisle`, `department`, etc. It contains the information of each order (e.g product name, number of product, the aisle it belongs to, the department it belongs to, etc.) .

1. How many aisles are there, and which aisles are the most items ordered from?
```{r aisle}
instacart |> 
  count(aisle) |> 
  arrange(desc(n))
```

There are 134 aisles in this dataset.
"fresh vegetables" are the most items ordered from.


2. Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered.
```{r}
instacart |> 
  count(aisle) |> 
  filter(n > 10000) |> 
  mutate(aisle = fct_reorder(aisle, n)) |> 
  ggplot(aes(x = aisle , y = n)) +
  geom_point() +
  labs( title = "Number of items ordered in each aisle") +
  theme(axis.text.x = element_text(angle = 70, hjust = 1))
```


3. Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.
```{r}
instacart |> 
  filter(aisle == c("baking ingredients", "dog food care", "packaged vegetables fruits")) |> 
  group_by(aisle) |> 
  count(product_name) |> 
  mutate(rank = min_rank(desc(n))) |> 
  filter(rank <= 3) |> 
  arrange(desc(n)) |> 
  knitr::kable()
```


4. Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).
```{r}
instacart |> 
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) |> 
  group_by(product_name, order_dow) |> 
  summarize(mean_hour = mean(order_hour_of_day)) |> 
  pivot_wider(
    names_from = order_dow,
    values_from = mean_hour
  ) |> 
  knitr::kable(digits = 2)
```


### Problem 2

load the data `brfss_smart2010`
```{r}
library(p8105.datasets)
data("brfss_smart2010")
```

Clean the data
```{r}
brfss_df =  
  brfss_smart2010 |> 
  janitor::clean_names() |> 
  filter(topic == "Overall Health") |> 
  filter(response %in% c("Excellent", "Very good", "Good", "Fair", "Poor")) |>
  mutate(response = forcats::fct_relevel(response, c("Poor", "Fair", "Good", "Very good", "Excellent")))
```


1. In 2002, which states were observed at 7 or more locations? What about in 2010?

```{r}
brfss_2002 =
  brfss_df |> 
  filter(year == 2002) |> 
  group_by(locationabbr) |> 
  summarize(n_location = n_distinct(locationdesc)) |> 
  filter(n_location >= 7)
brfss_2002
```

In 2002, 6 states: CT, FL, MA, MC, NJ, and PA were observed at 7 or more locations.

```{r}
brfss_2010 =
  brfss_df |> 
  filter(year == 2010) |> 
  group_by(locationabbr) |> 
  summarize(n_location = n_distinct(locationdesc)) |> 
  filter(n_location >= 7)
brfss_2010
```

In 2010, 14 states: CA, CO, FL, MA, MD, NC, NE, NJ, NY, OH, PA, SC, TX, WA were observed at 7 or more locations.

2. Construct a dataset that is limited to Excellent responses, and contains, year, state, and a variable that averages the data_value across locations within a state. Make a “spaghetti” plot of this average value over time within a state (that is, make a plot showing a line for each state across years – the geom_line geometry and group aesthetic will help).
```{r}
brfss_excellent =
  brfss_df |> 
  filter(response == "Excellent") |> 
  select(year, locationabbr, data_value) |> 
  group_by(year, locationabbr) |> 
  summarize(average_data = mean(data_value))
```

```{r spaghetti plot}
brfss_excellent |> 
  ggplot(aes(x = year, y = average_data, group = locationabbr)) + 
  geom_line(aes(color = locationabbr)) +
  labs(
    title = "Average Value of Excellent Responses Over Years",
    x = "Year",
    y = "Average Value",
    color = "state"
  )
```


3. Make a two-panel plot showing, for the years 2006, and 2010, distribution of data_value for responses (“Poor” to “Excellent”) among locations in NY State.

```{r}
brfss_df |> 
  filter(locationabbr == "NY") |> 
  filter(year %in% c(2006, 2010)) |> 
  ggplot(aes(x = response, y = data_value)) +
  geom_boxplot() +
  facet_grid( ~ year)
```



### Problem 3

load and tidy the paticipants data
```{r}
participants_df =
  read_csv("data/nhanes_covar.csv", skip = 4) |> 
  janitor::clean_names() |> 
  filter(age >= 21) |> 
  drop_na() |> 
  mutate(
    sex = case_match(
      sex,
      1 ~"male",
      2 ~"female"),
    education = case_match(
      education,
      1 ~ "less than high school",
      2 ~ "high school equivalent",
      3 ~ "more than high school"),
    sex = as.factor(sex),
    education = as.factor(education))
```

load and tidy the accelerometer data
```{r}
accelerometer_df =
  read_csv("data/nhanes_accel.csv") |> 
  janitor::clean_names() 
```

merge two datasets
```{r}
combined_df = 
  inner_join(participants_df, accelerometer_df, by = "seqn") 
combined_df
```

Produce a reader-friendly table for the number of men and women in each education category, and create a visualization of the age distributions for men and women in each education category. Comment on these items.

```{r table}
combined_df |> 
  group_by(education, sex) |> 
  summarize(n_sex = n()) |> 
  pivot_wider(
    names_from = sex,
    values_from = n_sex
  ) |> 
  knitr::kable()
```

```{r age distribution graph}
combined_df |> 
  ggplot(aes(x = education, y = age, color = sex)) +
  geom_boxplot() +
  labs(
    title = "Age Distribution for Education Level of Male and Female Participants",
    x = "Education level",
    y = "Age"
  )
```



Traditional analyses of accelerometer data focus on the total activity over the day. Using your tidied dataset, aggregate across minutes to create a total activity variable for each participant. 
```{r}
total_activity = 
  combined_df |> 
  group_by(seqn, sex, age, education) |> 
  pivot_longer(
    min1:min1440,
    names_to = "minute",
    values_to = "activity"
  ) |> 
  summarize(total_activity = sum(activity))
```

```{r}
total_activity |> 
  mutate(education = forcats::fct_relevel(education, c("less than high school", "high school equivalent", "more than high school"))) |> 
  ggplot(aes(x = age, y = total_activity, color = sex)) + 
  geom_point() + 
  geom_smooth(se = FALSE) +
  facet_grid(.~ education) +
  labs(
    title = "Total Activitt time for Each Participant",
    x = "Age",
    y = "Total Activity Time"
  )
```

In this plot, I can conclude that total activity time are becoming less and less as age increases whatever his/her education level is. And participants who have more than high school education level have the least activity time on average among the three education levels. Female with high school equivalent and more than high school education level have more activity time compared males, while vice versa in those who are less than high school.


Accelerometer data allows the inspection activity over the course of the day. Make a three-panel plot that shows the 24-hour activity time courses for each education level and use color to indicate sex. 
```{r}
min_df = 
  combined_df |> 
  pivot_longer(
    min1:min1440,
    names_to = "minute",
    names_prefix = "min",
    values_to = "activity"
  ) |> 
  mutate(minute = as.numeric(minute))
```


```{r}
min_df |> 
  ggplot(aes(x = minute, y = activity, color = sex)) +
  geom_point(alpha = .3) +
  geom_smooth(se = FALSE) +
  facet_grid(.~ education) +
  scale_x_continuous(
    breaks = c(0, 360, 720, 1080, 1440)
    )+
  labs(
    title = "24-hour Activity Time Courses For Each Education Level",
    x = "Time in Minute",
    y = "Activity"
  )+
  theme(legend.position = "bottom")
```

In this plot, I can conclude that, in general, people remain a low level of activity during 0-6 am, and a high level of activity after 6 am, this is because many people get up after 6 am. While many males who are more than high school will still be active during 9-10 pm compared to others.
